\documentclass{beamer}
\usetheme{metropolis} % Clean modern look

% \usepackage{fontspec} % Needed for Unicode, if using emojis
\usepackage{amsmath, amsfonts}
\usepackage{graphicx}
\usepackage{booktabs}

\title{Exploratory Data Analysis (EDA)}
\subtitle{Theory and Math Behind Key Concepts}
% \author{Your Name}
\date{}

\begin{document}
{
\setbeamertemplate{frame footer}{\href{https://github.com/abkafi1234/Datamining_MachineLearning_Class/tree/main/Class_Slides}{Report Error}}
\setbeamerfont{frame footer}{size=\tiny} 
\begin{frame}
    \titlepage
\end{frame}
}
\begin{frame}{Outline}
    \tableofcontents
\end{frame}

% --- Slide: Missing Values ---
\section{Missing Values}
\begin{frame}{Missing Values}
\textbf{Definition:} Data entries where values are not recorded. \medskip

\textbf{Types:}
\begin{itemize}
    \item MCAR – Missing Completely at Random
    \item MAR – Missing at Random
    \item MNAR – Missing Not at Random
\end{itemize}

\textbf{Imputation Example: Mean}
\[
x_{\text{imputed}} = \frac{1}{n} \sum_{i=1}^{n} x_i
\]
\end{frame}

% --- Slide: Outliers ---
\section{Outliers}
\begin{frame}{Outliers}
\textbf{Definition:} Observations significantly different from the rest.

\textbf{Detection Methods:}
\begin{itemize}
    \item \textbf{IQR Method:}
    \[
    \text{IQR} = Q_3 - Q_1,\quad 
    \text{Lower} = Q_1 - 1.5 \times \text{IQR},\quad
    \text{Upper} = Q_3 + 1.5 \times \text{IQR}
    \]
    \item \textbf{Z-Score Method:}
    \[
    z_i = \frac{x_i - \mu}{\sigma}
    \]
    \item If \( |z_i| > 3 \), it's an outlier.
\end{itemize}
\end{frame}

% --- Slide: Irrelevant / Redundant Data ---
\section{Irrelevant / Redundant Data}
\begin{frame}{Irrelevant or Redundant Data}
\textbf{Definition:} Features that add little or no useful information.

\textbf{Methods:}
\begin{itemize}
    \item Correlation Analysis
    \item Variance Thresholding
    \item Principal Component Analysis (PCA)
\end{itemize}

\textbf{Pearson Correlation:}
\[
\rho_{X,Y} = \frac{\text{Cov}(X, Y)}{\sigma_X \sigma_Y}
\]
\end{frame}

% --- Slide: Data Type Correction ---
\section{Data Type Correction}
\begin{frame}{Data Type Correction}
\textbf{Goal:} Ensure appropriate types for modeling (e.g., integers, floats, categories).

\textbf{Examples:}
\begin{itemize}
    \item Convert strings to \texttt{datetime}
    \item Map numerical codes to categories
    \item Cast floats to integers (if discrete)
\end{itemize}
\end{frame}

% --- Slide: Categorical Variables ---
\section{Categorical Variables}
\begin{frame}{Categorical Variables}
\textbf{Types:}
\begin{itemize}
    \item Nominal (e.g., color)
    \item Ordinal (e.g., education level)
\end{itemize}

\textbf{Encoding Methods:}
\begin{itemize}
    \item \textbf{One-Hot Encoding:} 
    \[
    \text{Red} \rightarrow [1, 0, 0],\quad \text{Blue} \rightarrow [0, 1, 0]
    \]
    \item \textbf{Label Encoding:} 
    \[
    \text{Low} = 0,\quad \text{Medium} = 1,\quad \text{High} = 2
    \]
\end{itemize}
\end{frame}

% --- Slide: Normalization ---
\section{Normalization and Standardization}
\begin{frame}{Normalization}
\textbf{Definition:} Rescale features to a fixed range, usually [0, 1].

\textbf{Formula:}
\[
x'_i = \frac{x_i - \min(x)}{\max(x) - \min(x)}
\]

\textbf{Use case:} Distance-based models like KNN, K-Means.
\end{frame}

% --- Slide: Standardization ---
\begin{frame}{Standardization}
\textbf{Definition:} Rescale features to mean 0 and standard deviation 1.

\textbf{Formula:}
\[
x'_i = \frac{x_i - \mu}{\sigma}
\]

\textbf{Use case:} Models assuming normal distribution (e.g., PCA, linear regression).
\end{frame}

% --- Slide: Summary Table ---
\section{Summary Table}
\begin{frame}{EDA Summary Table}
\small
\begin{tabular}{@{}lll@{}}
\toprule
\textbf{Step} & \textbf{Method} & \textbf{Math} \\ \midrule
Missing Values & Mean Imputation & \( x' = \frac{1}{n} \sum x_i \) \\
Outliers & IQR, Z-score & \( z = \frac{x - \mu}{\sigma} \) \\
Redundant Data & Correlation, PCA & \( \rho = \frac{\text{Cov}(X,Y)}{\sigma_X \sigma_Y} \) \\
Data Types & Type casting & N/A \\
Categorical & One-Hot/Label Encoding & Encoding to vectors or integers \\
Normalization & Min-Max Scaling & \( x' = \frac{x - \min(x)}{\max(x) - \min(x)} \) \\
Standardization & Z-score Scaling & \( x' = \frac{x - \mu}{\sigma} \) \\
\bottomrule
\end{tabular}
\end{frame}
\begin{frame}[standout]
    Thank you!
\end{frame}

\end{document}
